BivData Setup

### 2020-05-14 doc/totalSC.Rmd
extensive debugging and organizing the datascripts

load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#batch18
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#most recent

#these data generated by running "python script --, it should be re-run when new cells are added to the anon.csv file
SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)

#load BivData
load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_13.2.20.RData")

@@

#strain specific xlsx sheets

@@

```{r write.table, echo=FALSE, eval=FALSE}
#write out these cell's info so I can double check them.
#write.table(Total.SC.above.threshold, "~./MLH1repo/data/BivData/total_SC_above_threshold.csv", sep=",", row.names = FALSE)

#write.table(MLH1.merge.Skel, "~./MLH1repo/data/BivData/total_SC_below_threshold.csv", sep=",", row.names = FALSE)
```

>>>>

load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#most recent

#these data generated by running "python script --, it should be re-run when new cells are added to the anon.csv file
SC.skel = read.csv("~./MLH1repo/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)

#load BivData  
load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_13.2.20.RData")

@@

#strain specific xlsx sheets (why?)

@@

# write table

```{r write.table, echo=FALSE, eval=FALSE}

#add categories and levels

#above threshold is bad,
#write out these cell's info so I can double check them.
#write.table(Total.SC.above.threshold, "~./MLH1repo/data/BivData/total_SC_above_threshold.csv", sep=",", row.names = FALSE)

#write.xlsx2(strain.table.HQ_Fig, "Main_MLH1_table2.xlsx", sheetName = "Sheet1", 
#            col.names = TRUE, row.names = FALSE, append = FALSE)

#write.table(MLH1.merge.Skel, "~./MLH1repo/data/BivData/total_SC_below_threshold.csv", sep=",", row.names = FALSE)

#Below threshold is are the values to use

#write.xlsx2(MLH1.merge.Skel, "Total.SC_merge.MLH1.xlsx", sheetName = "Sheet1", 
#            col.names = TRUE, row.names = FALSE, append = FALSE)
```

### 2020-05-14 src/BivData_Setup.Rmd
extensive debugging and organizing the datascripts

@@

#write files for basic stats to csv files
#OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
#OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)

@@

```{r split.up.DF, echo=FALSE}
#new.Curate.FULL to Curated_BivData

@@

#don't think I need this one
#Foci_TRUE_Curate_BivData <-   Curated_BivData[(!is.na(Curated_BivData$hand.foci.count)),]
```

```{r set.type_Curated_BivData, echo=FALSE}
##change factor cols to numbers (factor -> chr -> number)
Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)

@@

Curated_BivData$Rec.group.exclu <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                                ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))

@@

```{r no.Foci.Pos, echo=FALSE}

#no posittion
#just.SC.length <- Curated_BivData[is.na(Curated_BivData$Foci1),] #NAs not registering
just.SC.length <- Curated_BivData[(Curated_BivData$Foci1 == ""),]

#table(just.SC.length$category)

#decide what to do with these lines
#how will they be represented in Mat Methods?
```

```{r input.check,echo=FALSE}

#10jan18_13jan17_LEW_f1_sp1_14_rev_12  '?'

##calq the nromalized.siscoten, double check the negative values;  #13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14, #27aug15_12jun15_WSB_m2_sp1_18_rev_18, #21nov17_16nov17_MSM_f3_sp1_3_rev_1

#foci.pass = 2  #5aug17_14jul17_MSM_f1_sp1_13_rev_3
#hand foci =0 but there is a foci position
#too high hand foci count

#check for mistakes that might have been made when inputing / curation process

#hand foci > 4
#foci pos > SC lengths
#negative PER.foci.pos

#it hink most (ALL) of these came from bug in my function (ugh)

#Curated_BivData <- add_IFD(Curated_BivData)
#negative.IFD <- Curated_BivData[which(Curated_BivData$IFD1_ABS < 0),]
#caught 2
#15mar16_16jan16_G_f1_sp1_4_rev
#27mar16_6mar16_G_m3_sp1_12_rev

#neg Nrm F1
#negative.IFD2 <- Curated_BivData[which(Curated_BivData$PER_Foci_1 < 0),]
#0 really?

```{r neg.IFD, echo=FALSE}

@@

#neg.IFD <- Curated_BivData[Curated_BivData$IFD1_ABS < 0, ]
#neg.ifd <- Curated_BivData %>% filter(hand.foci.count > 1) %>% filter(IFD3_ABS < 0 )

@@

```{r CLEAN.UP.1, echo=FALSE, include=FALSE}

dup.curated <- new.Curate.FULL[duplicated(new.Curate.FULL$Obj.ID),] #784
dup.curated <- add_mouse(dup.curated)

#table(dup.curated$mouse)
#1mar15_PWD_m1

```{r remove.mice, echo=FALSE, eval=FALSE}

#mice to remove from the Curated_BivData -- should be similar to the mice exclude from the MLH1

#poor quality?

#reaxamine these mice for if they should be included in 
exclude_mice_list_MLH1 <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                  "8jun15_G_f3", "8jun15_G_f4",
                  "4jan17_LEW_f2",
                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
                  "12sep16_MSM_f3","31aug16_MSM_m2",
                  "27oct18_MOLF_m3",
                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",

    "3jul18_TOM_m1","3sep18_TOM_m1",

                "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
                  )
#REaxamine and rename to exclude mice_BivData

exmice.Curated_BivData <- Curated_BivData[ Curated_BivData$mouse %in% exclude_mice_list, ]

#old males

@@

```{r WSB.SC, echo=FALSE}
#make the dataset
WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
#correct and remove, 26dec17_18nov17_WSB_f4_sp1_12_rev_28, 26dec17_18nov17_WSB_f4_sp1_12_rev_27
table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)

#plot the differences
wsb.ploty <- ggplot(WSB.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()
wsb.ploty

@@

```{r LEW.SC, echo=FALSE}
#make the dataset
LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)

#plot the differences
LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

@@

```{r G.SC, echo=FALSE}
#make the dataset
G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)

#plot the differences
G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

@@

```{r PWD.SC, echo=FALSE}
#make the dataset
PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)

#plot the differences
PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

@@

```{r KAZ.SC, echo=FALSE}
#make the dataset
KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)

#plot the differences
KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

@@

```{r SKIVE.Sc, echo=FALSE}
#make the dataset
SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)

#plot the differences
SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

@@

```{r MSM.SC, echo=FALSE}
#make the dataset
MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)

#plot the differences
MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

@@

```{r MOLF.SC, eval=FALSE}
#make the dataset
MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)

#plot the differences
MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")

@@

@@

# Long and Short BivData Set

@@

#1. Isolate bivalent measrues from 'almost whole cells'   (how many biv's are from cells with >15 bivalents)
over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]

@@

```{r short.biv.data, echo=FALSE, include=FALSE, eval=FALSE}

@@

                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

@@

#save Rdata
save.image("~./MLH1repo/data/CleanBivData_6.3.20.RData")
#save.image("data/BivData_4_PCA.RData")

>>>>

files that will be made in setup
  5. CELL level table (full all mice) (use this to make long biv dataset)

@@

#add_mouse ect.
#write files for basic stats to csv files
#OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
#OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)

@@

```{r mk.Curate_BivData, echo=FALSE}
#new.Curate.FULL to Curated_BivData

@@

#don't think I need this one
#Foci_TRUE_Curate_BivData <-   Curated_BivData[(!is.na(Curated_BivData$hand.foci.count)),]

##change factor cols to numbers (factor -> chr -> number)
Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)

@@

#remove bivalent entries without hand foci count values
cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )
```

@@

```{r no.Foci.Pos, echo=FALSE, eval=FALSE}

#no posittion
#just.SC.length <- Curated_BivData[is.na(Curated_BivData$Foci1),] #NAs not registering

#not sure why these are different DFs
just.SC.length2 <- Curated_BivData %>% filter(Foci1 == " ")
just.SC.length <- Curated_BivData[(Curated_BivData$Foci1 == ""),]
#these are different than hand.foci.count

table(Curated_BivData$hand.foci.count, exclude = NULL)
#there are ~600 rows with SC and no hand foci count (blah)
```

```{r neg.IFD, echo=FALSE, eval=FALSE}

@@

neg.IFD <- Curated_BivData[Curated_BivData$IFD1_ABS < 0, ]
neg.ifd <- Curated_BivData %>% filter(hand.foci.count > 1) %>% filter(IFD3_ABS < 0 )

#same deal as above
```

```{r remove.DUP, echo=FALSE}

dup.curated <- Curated_BivData[duplicated(Curated_BivData$Obj.ID),] #784
dup.curated <- add_mouse(dup.curated)

#table(dup.curated$category)#MSM male, PWD male
#obj.ID is not complete for all rows

Curated_BivData_w.dup <- Curated_BivData

Curated_BivData <- Curated_BivData %>% distinct(Obj.ID, .keep_all = TRUE)

Duplicates removed, `r length(Curated_BivData_w.dup$Obj.ID) - length(Curated_BivData$Obj.ID)`. most were MSM and PWD males.

```{r remove.mice, echo=FALSE, eval=TRUE}

#reaxamine these mice for if they should be included in 
#this list is the same as setupData
exclude_mice_list_MLH1 <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                  "8jun15_G_f3", "8jun15_G_f4",
                  "4jan17_LEW_f2",
                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
                  "12sep16_MSM_f3","31aug16_MSM_m2",
                  "27oct18_MOLF_m3",
                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",

    "3jul18_TOM_m1","3sep18_TOM_m1",
                "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
                  )
#REaxamine and rename to exclude mice_BivData

exmice.Curated_BivData <- Curated_BivData[ Curated_BivData$mouse %in% exclude_mice_list_MLH1, ]

Curated_BivData <- Curated_BivData[ ! Curated_BivData$mouse %in% exclude_mice_list_MLH1, ]

`r length(exmice.Curated_BivData$Obj.ID)` rows removed for excluded mice.

after all of these chunks are run, save the excel file for Curated_BivData

DUPs, negative IFDs, remove juvi and MLH1 exclude mice,

There are 3 bivalents with Per.Foci2 > 1 (remove or go correct)

```{r write.Curated.BivData, echo=FALSE}

#this file should have all the category columns, and pass the curation levels above

#write.xlsx2(Curated_BivData, "Curated_BivData.xlsx", sheetName = "Sheet1", 
#            col.names = TRUE, row.names = FALSE, append = FALSE)
#this excel causes an error

#write.csv(Curated_BivData, "Curated_BivData.csv",   row.names = FALSE)

@@

@@

```{r WSB.SC, echo=FALSE}
#make the dataset
WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
#correct and remove, 26dec17_18nov17_WSB_f4_sp1_12_rev_28, 26dec17_18nov17_WSB_f4_sp1_12_rev_27
table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count, exclude = NULL)

#plot the differences
wsb.ploty <- ggplot(WSB.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(WSB.Bivalents.DF$strain)
wsb.ploty

@@

)
#can I plot these stats
```

```{r LEW.SC, echo=FALSE}
#make the dataset
LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count, exclude = NULL)

#plot the differences
LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(LEW.Bivalents.DF$strain)

@@

```{r G.SC, echo=FALSE}
#make the dataset
G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count, exclude = NULL)

#plot the differences
G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(G.Bivalents.DF$strain)

@@

```{r PWD.SC, echo=FALSE}
#make the dataset
PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count, exclude = NULL)

#plot the differences
PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(PWD.Bivalents.DF$strain)

@@

```{r KAZ.SC, echo=FALSE}
#make the dataset
KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count, exclude = NULL)

#plot the differences
KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(KAZ.Bivalents.DF$strain)

@@

```{r SKIVE.Sc, echo=FALSE}
#make the dataset
SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count, exclude = NULL)

#plot the differences
SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(SKIVE.Bivalents.DF$strain)

@@

```{r MSM.SC, echo=FALSE}
#make the dataset
MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count, exclude = NULL)

#plot the differences
MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(MSM.Bivalents.DF$strain)

@@

```{r MOLF.SC, echo=FALSE, eval=FALSE}
#make the dataset
MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count, exclude = NULL)

#plot the differences
MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(MOLF.Bivalents.DF$strain)

@@

Redo the plots from above, just male data with strain on the x

```{r BivLengths_by_strain, echo=FALSE}

#set the order of strains

BivLen.Male <- ggplot(
  (Curated_BivData %>% filter(sex == "male") %>% filter(species == "M.musculus")),  aes(y=chromosomeLength, x = strain, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("Male Single Bivalent Lengths")
BivLen.Male
```

@@

# reduced Long and Short BivData Set

@@

#1. Isolate bivalent measrues from 'almost whole cells'   (how many biv's are from cells with >15 bivalents)

over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]

@@

```{r short.biv.data, echo=FALSE, include=FALSE, eval=TRUE}

@@

                        Nmice = length(unique(mouse)),
                        nobs  = length(Obj.ID),

@@

#save Rdata
save.image("~./MLH1repo/data/CleanBivData_5.12.20.RData")
#save.image("data/BivData_4_PCA.RData")
